Palaeodistribution of crop progenitors and other plant resources over the Pleistocene–Holocene transition in Southwest Asia

### 2023-09-01 analysis/analysis.qmd
Outline introductory sections

---
title: "Palaeodistribution of crop progenitors and wild foods in the Southern Levant"
author: "Joe Roe"
format: html
bibliography: references.bib
---

@@

@@

@@

>>>>

---
title: Palaeodistribution of crop progenitors and other plant resources over the Pleistocene–Holocene transition in Southwest Asia
date: today
author: 
- name: Joe Roe
  id: jr
  orcid: 0000-0002-1011-1244
  email: joeroe@hey.com
  affiliation: University of Bern
- name: Amaia Arranz-Otaegui
  id: aao
  orcid: 0000-0002-5091-6426
  affiliation: University of the Basque Country
abstract: |
  This paper presents the first comprehensive reconstructions of the prehistoric 
  ranges of [X] plant species known to have been gathered or cultivated by early
  agricultural societies in Southwest Asia, including the progenitors of the 
  first crops. We used machine learning to train an ecological niche model (ENM)
  of each species based on its present-day distribution in relation to climate 
  and environmental variables. Predictions of the potential ranges of these 
  species at key stages of the Pleistocene–Holocene transition could then be
  derived from these models using palaeoclimate 
  Species ranges were on average [X%] [larger|smaller] in the Early Holocene 
  compared to today, indicating [...]. The modelled ranges predict the
  observed occurrence of these species on archaeological sites with [low|medium|high]
  accuracy. The regional ubiquity of species in the archaeological record is [not] 
  correlated with the predicted size of its range and the diversity of
  archaeobotanical assemblages is [not] correlated with the predicted diversity
  of its environs. This indicates that trends in taxonomic composition of
  the archaeobotanical record is [not] likely to be influenced by environmental 
  change and species turnover, not just, as is often assumed, human economic 
  choices.
format: html
bibliography: references.bib
---

@@

* The first farming societies had an ecological context
* Subsistence is understood (largely) through archaeobotany; ecological context from archaeobotany, palynology, palaeoclimate, etc.
  * But these have a variety of biases (human selection, taphonomy, etc.)
  * And at the end of the day only represent specific places – interpolation to the entire region is not straightforward
  * Overlap makes it difficult to see where human choices depart from the environmental background [cf. Martin et al. 2016, for zooarch]
* Here we present an alternative approach using ENM
  * Whole region, at multiple climate snapshots
  * Independent of archaeobot. and pal. clim. data, so can verify and compare

@@

* Epipal./Neo. plant-based economies
  * More than the "founder crops"; 
  * More than food;
  * Regionally and temporally diverse
  * ...so we model lots of species!
* Regional ecological reconstructions
  * See CSEAS (AEA-prep) presentation
  * Figure: comparisons
* Archaeological applications of ENM
  * Review: Franklin et al. 2015
  * In SW Asia: Conolly et al. 2012, Collins et al. 2018
  * Train with archaeo, or train with present-day?
    * In ENM literature, many sceptical about fore- and hind-casting
    * Archaeo data a golden opportunity here

@@

## Acknowledgments {.appendix}
